---
title: "Intro to Quarto"
author: "Julia Lober"
format: html
bibliography: references.bib
---

# Big title

#### Basic math

Let's look at $2 \times 2$.

```{r}
#| echo: false
2 * 2
```

Since that is so simple, we don't really need to see the code.

[Google](https://google.com) could also have given us that answer pretty easily.

#### A list

-   we can write multiple things
-   in a row
-   like this.

```{r}
a <- 1:10 

for (num in a) {
  print(num * num)
}
```

## Citations!

Here is a fact [@oke2020] with a citation.

This is from the `Insert` menu in `Visual` edit mode. You can put a DOI into the search box and RStudio/Quarto will search for the materials and generate a citation for the reference pretty much automatically.

You *can* write manuscripts in Quarto. This would be something for YAML set-up.

## Table

| Col1    | Col2 | Another column |
|---------|------|----------------|
| Nothing | here | still          |

: Table 1. has a caption now.

------------------------------------------------------------------------

## Code chunks.

Create a new chunk with `Opt-Alt-I`

```{r}

```
