---
title: "Mini Literate Analysis"
author: "Julia Lober" 
format: html
---

## Introduction

### About the data

This data set contains chemical oceanographic data from the North Pole including CDT parameters and a variety of bottle measurements. This data is publicly available at the Arctic Data Center [here](https://arcticdata.io/catalog/view/doi:10.18739/A25T3FZ8X). 

### Setup

These are the packages that we will need for our analysis. 

```{r}
library(readr)
```

### Read data

We read in the data as a csv. 

```{r}
data <- read_csv("data/BGchem2008data.csv", show_col_types = FALSE)
print(data)
```

### Exploring data

```{r}
colnames(data)
summary(data)
unique(data$Date)
```


## Analysis

### Calculate summary statistics

```{r}
nitrate <- mean(data$NO3)
nitrite <- mean(data$NO2)
amm <- mean(data$NH4)
phos <- mean(data$P)
```

The average nitrate is `r round(nitrate, digits = 3)`, nitrite is `r round(nitrite, digits = 3)`, ammonium is `r round(amm, digits = 3)`, and phosphorous is `r round(phos, digits = 3)`.

### Calculate the mean Readfield ratio

The Readfield ratio is an expected ratio of Carbon:Nitrogen:Phosphorous in the worlds oceans, which is generally expected to be a constant 106:16:1.  

```{r}
readfield_ratio <- (nitrate + nitrite + amm) / phos
```

### Plot ratio

```{r}
plot(data$P, (data$NO2 + data$NO3 + data$NH4),
     xlab = "Phosphorous",
     ylab = "Nitrogen",
     main = "Readfield Ratio")
```


## Conclusion

The Readfield ratio for this dataset is `r round(readfield_ratio, digits = 3)`. 

